---
title: "Homework MD 06 extra"
author: "Logan Agin"
format: html
editor: visual
embed-resources: true
---

# [Homework 6.2](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#homework-62)

For this assignment you are going to use the `bikes` dataset we began analyzing in class. You can get that dataset from the **bayesrules** package. Once **bayesrules** is installed, you can begin.

```{r}
#| message: false
library(tidyverse)
library(moderndive)
theme_set(theme_minimal())
```

Load the data and take a look.

```{r}
data(bikes, package = "bayesrules") 
glimpse(bikes)
```

## [Question 0](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-0)

Before analyzing a dataset, it's good to know some stuff about it. Answer the following questions:

-   how many rows are in the dataset?

    500

-   what does each row represent?

    a day

-   what dates does the dataset cover?

    dates that people rode bikes

-   what is the highest observed ridership in the dataset?

    6946

-   what was the highest wind speed recorded in the dataset?

    \~34

## [Question 1](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-1)

What is the correlation between number of rides and what the temperature feels like (in Fahrenheit)? What is the correlation between the number of rides and wind speed (miles per hour)?

```{r}
mod1 <- lm(rides ~ temp_actual,
           data = bikes)
get_regression_table(mod1)

get_correlation(rides ~ temp_actual,
                data = bikes)

mod2 <- lm(rides~windspeed,
           data = bikes)
get_regression_table(mod2)
get_correlation(rides~windspeed,
                data = bikes)
```

The correlation temp_actual = .58

The correlation windspeed = -.195

## [Question 2](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-2)

Using the approximation that a mile is equal to 1.61 kilometers, convert `windspeed` to kilometers per hour. Call the new variable `wind_kph` and add it to the `bikes` data frame. What is the correlation between wind speed in MPH and wind speed in KPH? Explain why in enough detail that I know you understand.

```{r}
bikes <- bikes %>% 
  mutate(wind_kph = windspeed * 1.61)

mod2kph <- lm(rides ~ wind_kph,
              data = bikes)
get_regression_table(mod2kph)

get_correlation(rides ~ wind_kph,
                data = bikes)

cor(bikes$windspeed , bikes$wind_kph)
```

The correlation between `windspeed` and `wind_kph` is 1. That is because they are the same measurement but one is just scaled to \* 1.61 for every point not changing the linear relationship.

## [Question 3](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-3)

Estimate two simple regressions:

-   predict rides as a function of wind speed in MPH

    rides i = 4205 -56(windspeed i)

-   predict rides as a function of wind speed in KPH

    rides i = 4205 - 35(wind_khp i)

Use `get_regression_table()` or `broom::tidy()` to display the results. This will give you two sets of an intercept and a slope. If any coefficients are the same between models, explain why. If any coefficients are different between models, explain why. Make sure to give me enough detail to convince me you understand.

```{r}
get_regression_table(mod2)
get_regression_table(mod2kph)

```

Though the correlation = 1 between `windspeed` and `wind_kph` their slopes are not the same because of the units of measurement. For the coefficients, we see for both 4205 meaning at windspeed (MPH) at zero, rides = 4205, and windspeed (KPH) at zero rides = 4205. This is because they are still measuring windspeed no matter what unit used, the intercept at windspeed 0 = 4205. The slope is different because the unit of measurement is scaled by 1.61. If you were to multiply the KPH slope by 1.61 you would get the MPH slope.

## [Question 4](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-4)

Using the models from above, tell me what the predicted ridership would be if the wind is blowing at 20 KPH. What would the predicted ridership be if the wind is blowing at 20 MPH?

rides i = -2169 + 89(windspeed i)

= -2169 + 89(20)

rides i = 4205 - 35(wind_khp i)

= 4205 -35(20)

or

```{r}
library(ggeffects)

predictions <- ggpredict(mod2,

                         terms = "windspeed",
                         ci_level = NA) 

predictions2 <- ggpredict(mod2kph,
                          terms = "wind_kph",
                          ci_level = NA)

predictions
predictions2
```

For MPH at windspeed 20 = 3095

For KPH at windspeed 20 = 3515

## [Question 5](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-5)

Let's get more realistic and move to multiple regression. We're going to use `temp_feel` in addition to wind speed. But we don't want to use Fahrenheit. So make a new variable called `temp_c` that is a conversion of `temp_feel` to Celsius and add it to the `bikes` dataframe. (You may have to look up how to do this conversion.)

```{r}
bikes <- bikes %>% 
  mutate(temp_c = (temp_feel-32)*(5/9))
```

With this new variable, estimate a multiple regression that predicts rides as an additive function of `wind_kph` and `temp_c` (that is, no interaction). Interpret both of the slope coefficients using the following language:

"For a \_\_\_\_\_ increase in \_\_\_\_\_, the model expects a \_\_\_\_\_ increase in \_\_\_\_\_."

Now interpret the intercept.

```{r}
mod3 <- lm(rides~ wind_kph + temp_c,
           data = bikes)
get_regression_table(mod3)

```

rides i = 942 - 20 (wind_kph i) + 143(temp_c i)

For a 1 increase in wind_kph, the model expects a -20 decrease in rides.

For a 1 increase in Celsius, the model expects a 143 increase in rides.

## [Question 6](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-6)

Using the multiple regression you just estimated, give me ridership predictions for the following situations:

-   SITUATION 1: temp = 25C, wind = 15 KPH

    4226

-   SITUATION 2: temp = 15C, wind = 5 KPH

    2992

-   SITUATION 3: temp = 10C, wind = 40 KPH

    1581

You can do this manually (using R as a calculator) or you might find the following code useful. I'm not going to give you *all* the code, but this is a head start!

```{r}
pred_df <- tibble(   
  situation = 1:3,   
  temp_c = c(25, 15, 10),   
  wind_kph = c(15, 5, 40) )
solution <- predict(mod3,
                    newdata = pred_df)
pred_df <- bind_cols(pred_df,solution)
colnames(pred_df) <- c("Situation", "Temp_Celsius", "Wind_KPH","Predicted_Ridership" )
pred_df
```

## [Question 7](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-7)

Let's add another predictor into the mix. Estimate a new model that uses `weekend` in addition to the predictors already in the model. Display the model results. Interpret the coefficient on `weekend` using a complete sentence.

```{r}
mod4 <- lm(rides ~ wind_kph + temp_c + weekend,
           data = bikes)
get_regression_table(mod4)
```

If it is the weekend, the model expects a -714 decrease in rides.

## [Question 8](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-8)

If the temperature and the wind speed are average, what is the expected ridership for a weekend day? What is the expected ridership for a weekday? Show the code that gets you your answers.

```{r}
library(dplyr)
bikes <- bikes %>% 
  mutate(weekend = as.factor(weekend))
mod5 <- lm(rides~ wind_kph + temp_c + weekend,
           data = bikes)

pred_df2 <- tibble(situation = 1:2,
                   temp_c = c(mean(bikes$temp_c),
                              mean(bikes$temp_c)),
                   wind_kph = c(mean(bikes$wind_kph),
                                mean(bikes$wind_kph)),
                   weekend = c("TRUE","FALSE")) 
predictions3 <- ggpredict(mod5,
                          terms = c("weekend", "temp_c", "wind_kph"),
                          conditions = pred_df2,
                          ci_level = NA)
predictions3

                   
                  

```

The expected ridership would be 2971.

## [Question 9](https://github.com/vaiseys/dav-course/blob/main/homework/chapter_06_extra.md#question-9)

You can use `get_regression_points()` or `predict()` to see how the model did at predicting each individual value of `rides`. Use one of these functions to find the date with the *largest absolute residual*. That is, find the day where the model is most wrong about predicted ridership. Why is the model so wrong about this day? (There is a correct answer here.)

```{r}
residuals <- resid(mod5)


max_residual_index <- which.max(abs(residuals))


date_with_largest_residual <- bikes$date[max_residual_index]


print(date_with_largest_residual)


```

We did not account for weather_cat. The weather category was a 3 but the temperature, weekday and wind_kph were normal. Since in the model we did not account for weather_cat this would throw it off.
